---
title: "Plotsfor MERS"
author: "Trevor Eakes"
date: "April 25, 2016"
output: html_document
---

I need to create good, solid plots for the MERS journal so I am creating a nice code to do just that. This code will keep fonts the same, make sure the file is saved in the correct size, and ensure everything is well sized. 

Let's do this

```{r manipulatingdata}
Adiv.abioticC <- Adiv.abiotic[-c(97:99),]
Rich <- DiversityIndex[-c(186:200), "Richness"]
Adiv.abioticC <- cbind(Adiv.abioticC, m=TetameResults$m)
Adiv.abioticC <- Adiv.abioticC[which(Adiv.abioticC[,"depth..m."]==0),]
```

Color Plot function and parameters

```{r plotfunctions}
dev.off() #whipe the slate clean

#x, y, col, and name variables must all be imputed to the function in ""
S <- data[paste(S)]

Water.Masses <- clades[,2]
for(i in 1:length(Water.Masses)) {
  
  value <- Water.Masses[i]
  if(value==1) {
    value <- "Kuroshio"}
  if(value==2){
    value <-"Deep"
  }
  if(value==3) {
    value <- "Oyashio"}
  Water.Masses[i] <- value
  }

TScladeplot <- function(Adiv.abiotic, S, Theta, Gccom){
EU <- ggplot(Fillament, aes(x=S, y=Theta, colour=Richness)) + geom_point(size=3, alpha=.9, position=position_dodge(), stat="identity")

+scale_color_gradientn(colours=jet.colors(7), space="rgb", guide="colourbar")   #can add this for continuous variables
EU <- EU+labs(x=x, y=y)
return(EU)}
```

```{r plotfunctions}
#name must be in ""
require(RColorBrewer)
myColors <- c("#4DAF4A","#E41A1C","#377EB8" )
names(myColors) <- levels(Water.Masses)
colScale <- scale_colour_manual(name = "Water.Masses",values = myColors)

Plotformat<- function(plot) {
EU <- EU+ylab("Potential Temeprature")+xlab("Salinity")
EU <- EU+

EU <- EU +theme(axis.title.x = element_text(color="cadetblue", vjust=-0.35, size=12, hjust=-.005, face="bold"), axis.title.y = element_text(color="cadetblue" , vjust=0.35, size=12, face="bold"),
axis.text.x=element_text(size=12, vjust=0.5, hjust=-.005), axis.text.y=element_text(size=12, vjust=.05),
panel.background = element_rect(fill = "white"),
plot.background = element_rect(fill = "white"),
axis.title = element_text(face = "bold",size = rel(1)),
axis.title.y = element_text(angle=90,vjust =2),
axis.title.x = element_text(vjust = 2),
axis.text = element_text(), 
axis.line = element_line(colour="black"),
axis.ticks = element_line(size=5),
panel.grid.major = element_line(colour="#f0f0f0"),
panel.grid.minor = element_blank(),
legend.key = element_rect(colour = NA),
legend.position = "bottom",
legend.direction = "horizontal",
legend.key.size= unit(5, "mm"),
legend.margin = unit(1, "mm"),
legend.title =element_blank(),
legend.text=element_text(size=12),
#guide_legend(title="Water Masses")),
plot.margin=unit(c(10,10,10,10),"mm"),
strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
strip.text = element_text(face="bold")
)
EU <- EU+colScale
#saving the plot for the journal
ggsave("TSplotdouble.pdf", plot= EU, device = pdf, path = NULL, scale = 2, width = 81 , height = 190, units = "mm", dpi = 800, limitsize = TRUE)

#Journal page area is 169mm by 81mm
return(Finalplot)
}
```
geomtext default of 5 might be 5mm and the theme() size unit is point. 1 point is 1/72 inch=0.35mm, so 1 in geom_text() is 1mm, 1/0.35 =~ 14/5

Other guidlines: Graphs should be submitted at about double the intended final size. Lettering should be neither excessively small nor too large: a guideline size is that upper-case letters should be about 4 mm high, i.e. about 2 mm high after reduction to final printed size. Avoid very thin or very thick lines.

```{r plotfunctions}
EU <- TScladeplot(Adiv.abiotic, S, Theta, Gccom)
EU<- Plotformat(EU) 
EU

```

```{r dend}
require(graphics)
ggplot() + 
  geom_segment(data=segments(dendr), aes(x=x, y=y, xend=xend, yend=yend)) + 
  geom_text(data=label(dendr), aes(x=x, y=y, label=label, hjust=0), size=3) +
  coord_flip() + scale_y_reverse(expand=c(0.2, 0)) + 
  theme(axis.line.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.y=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_rect(fill="white"),
        panel.grid=element_blank())
```

```{r}
clus1 <- as.numeric(apply(dfcluster1[,4:ncol(dfcluster1)], 2, function(x) sum(x>0)))
clus2 <- as.numeric(apply(dfcluster2[,4:ncol(dfcluster2)], 2, function(x) sum(x>0)))
clus3 <- as.numeric(apply(dfcluster3[,4:ncol(dfcluster3)], 2, function(x) sum(x>0)))

clus1c <- round(clus1/104, 2)*100
clus2c <- round(clus2/25, 2)*100
clus3c <- round(clus3/56, 2)*100

Incidence <- rbind(Kuroshio=clus1c, Oyashio=clus3c, Deep=clus2c)
Kuroshio_Phytoplankton <- Kuroshio_Phytoplankton.csv[-c(186:190), -c(1,2)]
coldendr(Incidence)<-colnames(Kuroshio_Phytoplankton)
Incidence <- (as.matrix(Incidence))
Incidence<-Incidence[order(Incidence[,1],decreasing=T),]
#
IncidenceMatrix <- heatmap(t(Incidence), Rowv=NA, Colv=NA, scale="row",margins=c(10,5))
#
#########################################################
### C) Customizing and plotting the heat map
#########################################################

# creates a own color palette from red to green
require(stats)
my_palette <- colorRampPalette(c("white","blue", "yellow", "red", "purple"))(n = 599)

# (optional) defines the color breaks manually for a "skewed" color transition
col_breaks = c(seq(0, 1, length=100), 
               seq(1, 2, length=100),
               seq(2,6,length=100),              
               seq(6,12,length=100),
               seq(12, 18, length=100),
               seq(18, 30, length=100))   
dev.off()
g <- heatmap.2(Incidence,
               cellnote = Incidence, # same data set for cell labels
               #offsetCol = 0.5,
                main = "% of times present in observations",   # heat map title
               notecol="black",      # change font color of cell labels to black
               #density.info="none",  # turns off density plot inside color legend
               trace="none",      # turns off trace lines inside the heat map
               margins =c(8,12),     # widens margins around plot
               col=my_palette,       # use on color palette defined earlier 
               breaks=col_breaks,    # enable color transition at specified limits
               dendrogram="none",   # only draw a row dendrogram
               Colv=NA,              # turn off column clustering
               key=T,
               symkey=FALSE,
               keysize=1.5,
               #lmat=rbind(c(0, 3, 4), c(2,1,0 ) ), lwid=c(1.5, 4, 2 ),
               Rowv=FALSE,
)

heatmap.2(Incidence,dendrogram="col", cellnote=Incidence,
          notecol="black",col=redblue(256),scale="none",key=TRUE, keysize=1.5,
          density.info="none", trace="none", cexRow=0.7,cexCol=1.2)


```

To change the plot font:

#install.packages("extrafont");library(extrafont)
#font_import("Trebuchet MS")
#library(ggplot2)
#qplot(1:10)+theme(text=element_text(family="Trebuchet MS"))

other useful plotting modifications

#EU <- EU+xlim(0, 2.5)+ylim(0, 2.5)
#EU <- EU+geom_text(aes(label=station),hjust=.5, vjust=-1)
#EU <- EU+stat_contour()

putting plots together

#library("gridExtra")
#grid.arrange(xdensity, blankPlot, scatterPlot, ydensity, ncol=2, nrow=2, widths=c(4, 1.4), heights=c(1.4, 4))
