---
title: "Clustering_Rareification"
author: "Trevor Eakes"
date: "May 8, 2016"
output: html_document
---
##Rarefaction Methods for the Kuroshio

####Rarefaction is a method for comparing diversity between sample groups which may have different coverage, area, or even distributions. It estimates the relationship between how diversity such as richness or the 1st order Hill number accumulates as more individuals/samples are collected. Recent development of this package and the idea behind rarefaction now allow for using abundance in adition to frequency. Previously I could only use frequency with this idea which is statistically less powerful. 

Research on rarefaction conducted by:
Anne Chao
National Tsing Hua University
Ecological Statistics, biodiversity

####First to prepare the data
Note: You must have the dataframe creation Rmd in your working directory.
It is titled Diversity&DataframeConstruction.Rmd
```{r}
centitest.Clade <- cbind(clades[-c(97:99),], centitest) #adding the clade t centitest

#Clustering the Fillament manually by salinity and position
Adiv.abioticR <- cbind("rows"=c(1:nrow(Adiv.abiotic)), Adiv.abiotic) # add rows to keep track of things
Fillament <- Adiv.abioticR[which(Adiv.abioticR[, "DfromF"]>0),] #Take stations north of the front
Fillament <- Fillament[which(Fillament[,"S"]>33.9),] #Take stations Above 33.9 salinity
Fillament <- Fillament[which(Fillament["station"]!=2),]# must manually input single station number you want removed each time
Fillament <- Fillament[which(Fillament["station"]!=3),]
Fillament <- Fillament[which(Fillament["lat"]>35.99),]
Fillament <- Fillament[which(Fillament["station"]!=2),]
require(ggplot2)
EU <- ggplot(Fillament, aes(x=lon, y=lat, colour=Theta)) + geom_point(size=10, position=position_dodge(), stat="identity")+scale_color_gradientn(colours=jet.colors(7), space="rgb", guide="colourbar")

fillament <- centitest.Clade[Fillament$rows,]
centitest.Clade <- centitest.Clade[-Fillament$rows,]
SplitData(centitest.Clade, 2, 'dfcluster')
#sums species in each cluster
clus1 <- as.numeric(apply(dfcluster1[,5:ncol(dfcluster1)], 2, function(x) sum(x>0)))
#gets rid of the 0's
clus1 <- c(nrow(dfcluster1), clus1[clus1>0])
clus2 <- as.numeric(apply(dfcluster2[,5:ncol(dfcluster2)], 2, function(x) sum(x>0)))
clus2 <- c(nrow(dfcluster2), clus2[clus2>0])
clus3 <- as.numeric(apply(dfcluster3[,5:ncol(dfcluster3)], 2, function(x) sum(x>0)))
clus3 <- c(nrow(dfcluster3), clus3[clus3>0])
clus4 <- as.numeric(apply(fillament[,5:ncol(fillament)], 2, function(x) sum(x>0)))
clus4 <- c(nrow(fillament), clus4[clus4>0])

#For abundance
clusA1 <- as.numeric(apply(dfcluster1[,5:ncol(dfcluster1)], 2, function(x) sum(x)))
clusA2 <- as.numeric(apply(dfcluster2[,5:ncol(dfcluster2)], 2, function(x) sum(x)))
clusA3 <- as.numeric(apply(dfcluster3[,5:ncol(dfcluster3)], 2, function(x) sum(x)))
clusA4 <- as.numeric(apply(fillament[,5:ncol(fillament)], 2, function(x) sum(x)))

#Making Rarefaction lists for the clusters with row names
Rarefaction <- list("Kuroshio"=clus1, "Fillament"=clus4, "Oyashio"=clus2, "Deep"=clus3)# "5"=clus5) 
RarefactionAbund <- list("Kuroshio"=clusA1, "Fillament"=clusA4, "Oyashio"=clusA2, "Deep"=clusA3)# "5"=clus5)
#--------------------------------------
require(iNEXT)
A0 <- iNEXT(RarefactionAbund, q=0,  datatype="abundance") #z is object class iNEXT. It provides all the data shown in the rarefaction graphs
A1 <- iNEXT(RarefactionAbund, q=1,  datatype="abundance") #z is object class iNEXT. It provides all the

z0 <- iNEXT(Rarefaction, q=0,  datatype="Incidence_freq") #z is object class iNEXT. It provides all the data shown in the rarefaction graphs
z1 <- iNEXT(Rarefaction, q=1,  datatype="incidence_freq") #z is object class iNEXT. It provides all the data 
RarefactionRichness <- ggiNEXT(A0, type = 1, facet.var = "none", color.var = "site") 
RarefactionRichness+ylab("Richness")
#This will be the plot I use for my talks
RarefactionH <- ggiNEXT(A1, type = 1, facet.var = "none", color.var = "site") 
RarefactionH+ylab("1st order Hill #")
RarefactionCoverage <- ggiNEXT(A1, type = 2, facet.var = "none", color.var = "site")
RarefactionCoverage+ylab("Sample coverage H1")
#
#Creates rarefaction graphs
```


```{r}
library(gplots)  # for heatmap.2
library(Heatplus)
library(phyloseq)
library(RColorBrewer)# for better color options

#rounding count data
#For rounding frequency count data and standardizing it to the sample size
clus1c <- round(clus1/104, 2)*100
clus2c <- round(clus2/25, 2)*100
clus3c <- round(clus3/56, 2)*100

Incidence <- data.frame(cbind(Kuroshio=clus1c, Oyashio=clus3c, Deep=clus2c)) #combining into single data frame

Kuroshio_Phytoplankton <- Kuroshio_Phytoplankton.csv[-c(97:99, 186:190), -c(1,2)] #Editing base data frame, getting rid of uneaded columns and rows
HeatAbund <- data.frame(RarefactionAbund)
#changing to a dataframe
row.names(HeatAbund)<-colnames(Kuroshio_Phytoplankton) #changing row names of Incidence to their true Phytoplankton names
HeatAbund<-HeatAbund[order(HeatAbund[,1],decreasing=TRUE),] #ordering rows from greatest to smallest
Rownames <- rownames(Incidence)
Colnames <- cbind("Kuroshio","Fillament", "Oyashio", "Deep")

col <-jet.colors(448)
col <-c("white", col)
# (optional) defines the color breaks manually for a "skewed" color transition
col_breaks = c(seq(0, 1, length=50), 
               seq(1, 2, length=50),
               seq(2,10,length=50),              
               seq(10,30,length=100),
               seq(30, 100, length=100),
               seq(100, 200, length=100))   
dev.off()
g <- heatmap.2(HeatAbund,
               cellnote = HeatAbund, # same data set for cell labels
               #offsetCol = 0.5,
               notecol="black",      # change font color of cell labels to black
               #density.info="none",  # turns off density plot inside color legend
               trace="none",      # turns off trace lines inside the heat map
               margins =c(8,12),     # widens margins around plot
               col=col,       # use on color palette defined earlier 
               breaks= col_breaks,    # enable color transition at specified limits
               dendrogram="none",   # only draw a row dendrogram
               Colv=NA,              # turn off column clustering
               key=TRUE,
               densadj = 0.25,
               symkey=TRUE,
               keysize=1.5,
               #lmat=rbind(c(0, 3, 4), c(2,1,0 ) ), lwid=c(1.5, 4, 2 ),
               Rowv=FALSE,
)



require(scales)
heat <- heat+scale_fill_continuous(low="lightblue", high="#FF3300", limits=c(1,100),breaks=c(2,5, 10, 15, 20, 30, 35), na.value = "white", guide_legend(title = "% Presence", keywidth = 1, keyheight = 10))

#p <- p+scale_fill_gradient2(low="lightblue", mid="yellow", high="#FF3300", limits=c(1,38),breaks=c(1, 2,5, 10, 15, 20, 30, 35), na.value = "white", guide_legend(title = "% Presence", keywidth = 1, keyheight = 3))
heat <- heat+xlab("")
heat <- heat+ylab("Species")
heat <- heat+theme(axis.title.x = element_text(color="cadetblue",hjust=1, vjust=.5, size=14, face="bold"), axis.title.y = element_text(color="cadetblue" , vjust=0.35, size=12, face="bold"),
axis.text.x=element_text(size=12, angle=330, face="bold"),
axis.text.y=element_text(size=10, vjust=.5, hjust=1), 
axis.text.y=element_text(size=12, vjust=.05),
panel.background = element_rect(fill = "white"),
plot.background = element_rect(fill = "white"),
legend.text=element_text(size=14),
legend.title=element_text(size=12, face="bold"),
legend.key.height= unit(35, "mm"),
legend.key.width = unit(8, "mm"),
#plot.margin=unit(c(10,10,10,10),"mm"),
strip.text = element_text(face="bold"))
heat
ggsave("heatmapnormal.pdf", plot= heat, device = "pdf", path = NULL, scale = 1, width = 50 , height = 100, units = "mm", dpi = 600, limitsize = TRUE)
```


Plotting the heat map using phyloseq
HeatAbund <- phyloseq(otu_table(HeatAbund, taxa_are_rows=TRUE)) #converts data to the correct formal
heat <- plot_heatmap(HeatAbund, method=NULL, na.value = "white", taxa_are_rows=TRUE,sample.order=Colnames, taxa.order= Rownames, low="lightblue", high="#FF3300", max.label=175)
#heat